Predicting I/O-performance in HPC using Artificial Neural Networks

### 2016-06-04 Poster/main.tex
Poster Endversion

Tools are demanded that help users of HPC-facilities to implement efficient input/output (I/O) in their programs.
It is difficult to find the best access parameters and patterns, due to complex parallel storage systems.
%Currently users have to optimize their programs at great expense to each system individually without much assistance.
To develop tools, which support the implementation of efficient I/O, a computational model of the storage system is key.\medskip

@@

	The processing of a file access in the storage system can be viewed using the I/O-path, which is \textbf{the path from the invoking processor to the storage medium that contains the data}.
	The resulting access time of the process depends on the depth of this I/O-path because storage media further along the path are increasingly slower.
	%the further the path leads into the storage system, the slower the storage media gets.
	While the first storage levels (Caches) are very quick to respond, the main memory is already magnitudes slower; the same applies for the step into the parallel storage system, that is connected via network to the computer nodes. %\medskip

@@

	The measurements were done on the super computer Mistral of the DKRZ (Deutsches Klimarechenzentrum).
	Mistral operates with the parallel distributed file system Lustre and consist of over 1500 computing nodes and 30 petabyte storage.
	\textbf{Two different use cases were tested: Sequential and random file access.}
	Access sizes varied from 1\,B to 16\,MiB.\medskip

	Our measurements have shown, as can be seen in figure \ref{exploration}, that the access times increase with access sizes.
	However, a phenomenon that can be explained with the I/O-path model can be seen as well: \textbf{Access times of measurements with equal parameter values are split into several groups of access times.}
	The file accesses can be distinguished by a step in the magnitude of access time that is caused by the different I/O-paths the system used.
	%Different I/O-paths were used for file accesses in these groups, which led to a step in the magnitude of the access time.

@@

The idea of error classes is based on the fact that \textbf{the residues of a simple model are characteristic for the groups of measurements with different typical access times} we obeserved in figure \ref{exploration}. The residues therefore represent the I/O-paths used by the storage system for the corresponding file accesses. 
%Here a simple model is a model that doesn't discriminate measurements with similar parameters.
Error classes are obtained by clustering the residues with a k-Means algorithm.

@@

	From the residues of our models (see table \ref{residues}) it became clear that access times shouldn't be modelled linearly.
	Furthermore, it has been found to be difficult to exploit time dependencies of the storage system.
	As our hypothesis predicted, \textbf{it is essential to use knowledge about I/O-paths for precise predictions of access times}.
	The error classes allow the ANN to decrease the error significantly.
	In figure \ref{preds} the benefit of error classes as input can be seen:
	The second ANN is often able to classify file accesses correctly in respect to their I/O-path.

	\begin{table}
		\scriptsize
		\vspace{0.2cm}
		\begin{tabular}{l|r|r|r|r}%
			&  \multicolumn{2}{|c}{MAE (s)}&  \multicolumn{2}{|c}{MSPE (\%)}\\ \hline
			%\toprule
			Modell & seq. access & rnd. access & seq. access & rnd. access\\ \hline
			%\midrule
			Linear regression & $7.6\cdot 10^-5$ & $4.76\cdot 10^-3$ & 59 & 14185  \\
			ANN & $6.0\cdot 10^-5$ & $3.13\cdot 10^-3$ & 22 & 530 \\
			ANN + time dependencies & $5.7\cdot 10^-5$ & $3.05\cdot 10^-3$ & 22 & 619\\
			ANN + error classes & $2.0\cdot 10^-5$ & $1.03\cdot 10^-3$ & 14 & 119\\
			%\bottomrule
		\end{tabular}
		\caption{Mean absolut errors and mean square percentage errors of the models for the two different file access patterns.}
		\label{residues}
	\end{table}

>>>>

Tools are demanded that help users of HPC-facilities to implement efficient input/output (I/O) in their programs.
It is difficult to find the best access parameters and patterns due to complex parallel storage systems.
%Currently users have to optimize their programs at great expense to each system individually without much assistance.
To develop tools which support the implementation of efficient I/O a computational model of the storage system is key.\medskip

@@

	The processing of a file access in the storage system can be viewed using the I/O-path which is \textbf{the path from the invoking processor to the storage medium that contains the data}.
	The resulting access time depends on the depth of this I/O-path because storage media further along the path are increasingly slower.
	%the further the path leads into the storage system, the slower the storage media gets.
	While the first storage levels (Caches) are very quick to respond, the main memory is already magnitudes slower; the same applies for the step into the parallel storage system, that is connected via network to the computer nodes. %\medskip

@@

	The measurements were done on the super computer Mistral of the DKRZ (Deutsches Klimarechenzentrum).
	Mistral operates with the parallel distributed file system Lustre and consists of over 1500 computing nodes and 30 petabyte storage.
	\textbf{Two different use cases were tested: Sequential and random file access.}
	Access sizes varied from 1\,B to 16\,MiB.\medskip

	Our measurements have shown, as can be seen in figure \ref{exploration}, that the access times increase with access sizes.
	However, a phenomenon that can be explained with the I/O-path model can be seen as well: \textbf{Access times of measurements with equal parameter values are split into several groups.}
	The file accesses can be distinguished by a step in the magnitude of access time that is caused by the different I/O-paths the system used.
	%Different I/O-paths were used for file accesses in these groups, which led to a step in the magnitude of the access time.

@@

The idea of error classes is based on the fact that \textbf{the residues of a simple model are characteristic for the groups of measurements with different typical access times} we observed in figure \ref{exploration}. The residues therefore represent the I/O-paths used by the storage system for the corresponding file accesses. 
%Here a simple model is a model that doesn't discriminate measurements with similar parameters.
Error classes are obtained by clustering the residues with a k-Means algorithm.

@@

	From the residues of our models (see table \ref{residues}) it became clear that access times shouldn't be modeled linearly.
	Furthermore, it has been found to be difficult to exploit time dependencies of the storage system.
	As our hypothesis predicted, \textbf{it is essential to use knowledge about I/O-paths for precise predictions of access times}.
	The error classes allow the ANN to decrease the error significantly.
	In figure \ref{preds} the benefit of error classes as input can be seen:
	The second ANN is often able to classify file accesses correctly in respect to their I/O-path.

	\begin{table}
		\scriptsize
		\vspace{0.2cm}
		\begin{tabular}{l|r|r|r|r}%
			&  \multicolumn{2}{|c}{MAE (s)}&  \multicolumn{2}{|c}{MSPE (\%)}\\ \hline
			%\toprule
			Model & seq. access & rnd. access & seq. access & rnd. access\\ \hline
			%\midrule
			Linear regression & $7.6\cdot 10^-5$ & $4.76\cdot 10^-3$ & 59 & 14185  \\
			ANN & $6.0\cdot 10^-5$ & $3.13\cdot 10^-3$ & 22 & 530 \\
			ANN + time dependencies & $5.7\cdot 10^-5$ & $3.05\cdot 10^-3$ & 22 & 619\\
			ANN + error classes & $2.0\cdot 10^-5$ & $1.03\cdot 10^-3$ & 14 & 119\\
			%\bottomrule
		\end{tabular}
		\caption{Mean absolute errors and mean square percentage errors of the models for the two different file access patterns.}
		\label{residues}
	\end{table}
